### 2021-11-07 paper/output/document/outline_whykant.tex
did goals
idk man

The argument is not that no form of consequentialism or virtue ethics is tractable to automate, but 

@@

of ethical theories, but instead sketch the basic principles that most consequentialist and virtue 
ethical theories share.%

@@

the three formulations of the categorical imperative, focusing on the Formula of Universal Law (FUL) \cite{groundwork}. 
This is not a full defense of Kantian ethics but is instead a quick sketch of one argument for the FUL.%

@@

The FUL is easy to formalize
because it is a purely formal principle that is only evaluates the form of the maxim that an agent 
acts on. No other information is relevant to the test, so moral judgement can proceed with a relatively small
amount of data.%

@@

maxim. The stances I take are generally accepted by most Kantians, though some still disagree.%

@@

I define
a consequentialist theory as one that evaluates the consequences of an action, acknowledging that
this definition itself is controversial \cite{consequentialismsep}. I then present the debates over theories of the good, which 
consequences to evaluate, and the aggregation of consequences.%

@@

and the complexity of making judgements about an entire state of affairs, as opposed to about a single person.
A reasoner will need large, complicated datasets to settle these questions.%

@@

I understand the concept of virtue as those traits that are 
good for the posessor \cite{vesep}. I briefly explain Aristotle's eudaimonistic conception of virtue and present
some examples of virtues (courage, temperance, equanimity).%

@@

Automated virtue ethics will need to plant a flag in the messy, controversial debate over the exact
list of virtues. While most Kantians agree on one interpretation of the FUL, most virtue ethicists 

@@

\isamarkupsubsubsection{Machine Learning and Virtue Ethics%

@@

There is a connection between the ideas of cultivating habit and mimicking virtuous action
and machine learning, which mimics patterns in datasets\footnote{Would love feedback on whether this should
go here or in a Related Work section.}. There's been some
work using machine learning to learn moral behavior from a dataset of actions tagged as ethical \cite{delphi}. 
One drawback of this approach is the fact that machine learning algorithms have trouble
explaining why they made the judgements they made and often pick up on patterns that human beings would 
not see as indicative of causation. In contrast,
my system can explain exactly which axioms and principles resulted in a maxim being obligated or prohibited
and can even present human-reconstructable proofs of its results.%

@@

%:%75=28%:%
%:%84=31%:%

@@

%:%98=35%:%
%:%99=36%:%
%:%108=39%:%
%:%120=41%:%
%:%121=42%:%
%:%122=43%:%
%:%123=44%:%
%:%132=46%:%
%:%136=48%:%
%:%148=50%:%
%:%149=51%:%
%:%150=52%:%
%:%151=53%:%
%:%160=55%:%
%:%164=57%:%
%:%176=59%:%
%:%177=60%:%
%:%178=61%:%
%:%179=62%:%
%:%188=64%:%
%:%200=66%:%
%:%201=67%:%
%:%202=68%:%
%:%203=69%:%
%:%204=70%:%
%:%213=72%:%
%:%217=74%:%
%:%229=76%:%
%:%230=77%:%
%:%231=78%:%
%:%240=80%:%
%:%244=82%:%
%:%256=85%:%
%:%257=86%:%
%:%258=87%:%
%:%267=89%:%
%:%279=91%:%
%:%280=92%:%
%:%289=94%:%
%:%301=96%:%
%:%302=97%:%
%:%303=98%:%
%:%304=99%:%
%:%305=100%:%
%:%306=101%:%
%:%307=102%:%
%:%308=103%:%
%:%309=104%:%

>>>>

I do not argue that consequentialism and virtue ethics are tractable to automate, but 

@@

of ethical theories, but instead sketch basic principles of consequentialist and virtue 
ethical theories.%

@@

the three formulations of the categorical imperative, focusing on the Formula of Universal Law (FUL) \cite{groundwork}.%

@@

The FUL is easy to formalize because it only evaluates the 
form of the maxim that an agent acts on. No other information is relevant to the test, so moral 
judgement can proceed with a relatively small amount of data.%

@@

maxim. The stances I take are generally accepted by most Kantians \cite{ebelsduggan, KorsgaardFUL}.%

@@

I define a consequentialist theory as one that evaluates the consequences of an action, acknowledging 
that this definition itself is controversial \cite{consequentialismsep}. I then present the debates 
over theories of the good, which consequences to evaluate, and the aggregation of consequences.%

@@

and the complexity of making judgements about an entire state of affairs, as opposed to about a single person.%
\isamarkupsubsubsection{Prior Work%
I briefly survey some prior work in formalizing consequentialism and argue that this work either
implicitly or explicitly takes stances on the debates above, resulting in oversimplification or 
punting the problem \cite{util1, util2}.%

@@

I understand the concept of virtue as those traits that are good for the posessor \cite{vesep}. 
I briefly explain Aristotle's eudaimonistic conception of virtue and present some examples of virtues 
(courage, temperance, equanimity).%

@@

Automated virtue ethics will need to plant a flag in the messy, controversial debate over the 
exact list of virtues. While most Kantians agree on one interpretation of the FUL, most virtue ethicists 

@@

\isamarkupsubsubsection{Prior Work: Machine Learning and Virtue Ethics%

@@

One argument for formalizing virtue ethics is the connection between cultivating habit
and machine learning, which learns patterns in datasets. There's been some work using machine learning 
to learn moral behavior from a dataset of actions tagged as ethical, but this approach takes an
implicit stance on the debate above \cite{delphi}. Moreover, machine learning algorithms have trouble
explaining why they made the judgements they made \cite{puiutta}. In contrast, my system can explain 
exactly which axioms resulted in a maxim being obligated and can even present human-reconstructable proofs.
I argue that explanability is particularly important for ethical judgements because these judgements 
are often controversial and high stakes.\footnote{There also seems to be something about Kantian ethics
that requires the will have a reason for adopting a maxim, for otherwise it would cede its autonomy
to the computer. I will flesh this out in a separate section that asks whether it's even a good idea 
to automate ethics.}%

@@

%:%83=30%:%
%:%95=32%:%

@@

%:%106=36%:%
%:%118=38%:%
%:%119=39%:%
%:%120=40%:%
%:%121=41%:%
%:%130=43%:%
%:%134=45%:%
%:%146=47%:%
%:%147=48%:%
%:%148=49%:%
%:%157=51%:%
%:%161=53%:%
%:%173=55%:%
%:%174=56%:%
%:%175=57%:%
%:%176=58%:%
%:%185=60%:%
%:%197=62%:%
%:%198=63%:%
%:%199=64%:%
%:%200=65%:%
%:%209=68%:%
%:%221=70%:%
%:%222=71%:%
%:%223=72%:%
%:%232=74%:%
%:%236=76%:%
%:%248=78%:%
%:%249=79%:%
%:%250=80%:%
%:%259=82%:%
%:%263=84%:%
%:%275=86%:%
%:%276=87%:%
%:%277=88%:%
%:%286=90%:%
%:%298=92%:%
%:%299=93%:%
%:%308=95%:%
%:%320=97%:%
%:%321=98%:%
%:%322=99%:%
%:%323=100%:%
%:%324=101%:%
%:%325=102%:%
%:%326=103%:%
%:%327=104%:%
%:%328=105%:%
%:%329=106%:%
%:%330=107%:%

### 2021-11-07 paper/output/document/paper.tex
some restructuring, added some morethoughts on maxims
jkldfa
idk man

%
\begin{isabellebody}%
\setisabellecontext{paper}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Introduction%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
As artifical reasoners become increasingly powerful, computers become capable of
performing complex ethical reasoning. The field of machine ethics \cite{mesurvey} is interesting for two reasons.
First, the proliferation of artifically autonomous agents is creating and will continue to create a demand for 
automated ethics. These agents must be able to reason about complex ethical theories 
that withstand philosophical scrutiny. Second, just as automated mathematical reasoning gives mathematicians
new powers, automated ethical reasoning is a tool that philosophers can use 
when reasoning about ethics. Many contradictions or paradoxes with an ethical system may not be 
 obvious to the human eye, but can be easily tested using an automated theorem prover.

Modelling ethics without sacrificing the intracacies of an ethical theory is a 
challenging computational and philosophical problem. Simple and intuitive computational approaches, 
such as encoding ethics as a constraint satisfaction problem, fail to capture
the complexity of most philosophically plausible systems. On the other hand, it is not immediately clear
how to formalize many complex moral theories, like virtue ethics.

Computational ethics also requires a sophisticated 
ethical theory to model. Constraint satisfaction systems often default to some version of utiliarianism, 
the principle of doing the most good for the most people. Alternatively, they model basic moral 
principles such as ``do not kill," without modelling the theory that these principles originated from.
Modelling a more complex ethical theory will not only enable smarter philosophical machines, it will
also empower philosophers to study more complex ethical issues with the computer's help. The entire
field of philosophy is devoted to developing and testing robust ethical theories. Plausible machine
ethics must draw on plausible moral philosophy.

The ideal candidate ethical theory will be both philosophically interesting and easy to 
formalize. Kantian ethics, often described as ``formal," has been often floated as such a candidate \cite{powers, BL, lin}. 
The categorical imperative, Kant's universal law of morality, is a moral rule that can be used to 
guide action.

This project's objective is to automate Kantian ethics. I present two different
formalizations of Kant's categorical imperative implemented and tested in the Isabelle/HOL \cite{isabelle} theorem prover.
I model each formalization  as an extension of Carmo and Jones' Dyadic Deontic Logic (DDL) \cite{CJDDL}. I then embed the corresponding 
DDL formalization in higher-order logic (HOL) and implement it in Isabelle. Section 
\ref{sec:naive} implements and tests the naive formalization, a ``control group" that is clearly implausible but 
demonstrates the methods and tools of my approach. Section \ref{sec:kroy} implements a more sophisticated 
formalization inspired by Moshe Kroy's partial formalization of the categorical imperative.

I contribute implementations of two different interpretations of the categorical imperative, 
examples of how each implementation can be used to model and solve ethical scenarios, and tests that
examine ethical and logical properties of the system, including logical consistency, consistency
of obligation, and possibility of permissibility. The implementations themselves are usable models 
of ethical principles and the tests represent the kind of philosophical work that formalized ethics 
can contribute.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{System Overview \label{sec:overview}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
The goal of this project is to automate sophisticated ethical reasoning. This requires three
components. First, the choice of an ethical theory that is both intuitively attractive and 
lends itself to formalization. Second, the choice of formal logic to model the theory in. Third, the 
choice of automation engine to implement the formal model in. Section \ref{sec:kant} introduces Kantian
ethics, Section \ref{sec:DDL} explains Carmo and Jones's Dyadic Deontic Logic \cite{CJDDL} as a 
base logic, and Section \ref{sec:isabelle} presents the Isabelle/HOL implementation of the logic.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Kantian Ethics \label{sec:kant}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Kantian ethics is an attractive choice of ethical theory to automates. Kant's writings inspired much of Western analytic philosophy. 
In addition to the rich neo-Kantian program, almost all major philosophical traditions after Kant have
engaged with his work. Much of Western libertarian political thought is inspired by Kant's deontology,
and his ethics have bled into household ethical thought. Deontology is seen as one of the three major 
schools of Western analytic ethics.

@@

For more of fine-grained properties of DDL see \cite{CJDDL} or this project's source code. DDL is a heavy logic and contains modal operators 
that aren't necessary for my analysis. While this expressivity is powerful, it may also cause performance
impacts. DDL has a large set of axioms involving quantification over complex higher-order
logical expressions. Proofs involving these axioms will be computationally expensive.  Benzmueller 
and Parent warned me that this may become a problem if Isabelle's automated proof tools begin to time out.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/Desktop/cs91r/paper/paper.thy%:%
%:%24=6%:%
%:%36=8%:%
%:%37=9%:%
%:%38=10%:%
%:%39=11%:%
%:%40=12%:%
%:%41=13%:%
%:%42=14%:%
%:%43=15%:%
%:%44=16%:%
%:%45=17%:%
%:%46=18%:%
%:%47=19%:%
%:%48=20%:%
%:%49=21%:%
%:%50=22%:%
%:%51=23%:%
%:%52=24%:%
%:%53=25%:%
%:%54=26%:%
%:%55=27%:%
%:%56=28%:%
%:%57=29%:%
%:%58=30%:%
%:%59=31%:%
%:%60=32%:%
%:%61=33%:%
%:%62=34%:%
%:%63=35%:%
%:%64=36%:%
%:%65=37%:%
%:%66=38%:%
%:%67=39%:%
%:%68=40%:%
%:%69=41%:%
%:%70=42%:%
%:%71=43%:%
%:%72=44%:%
%:%73=45%:%
%:%74=46%:%
%:%75=47%:%
%:%76=48%:%
%:%77=49%:%
%:%78=50%:%
%:%87=52%:%
%:%99=54%:%
%:%100=55%:%
%:%101=56%:%
%:%102=57%:%
%:%103=58%:%
%:%104=59%:%
%:%113=61%:%
%:%125=63%:%
%:%126=64%:%
%:%127=65%:%
%:%128=66%:%
%:%129=67%:%
%:%130=68%:%
%:%131=69%:%
%:%132=70%:%
%:%133=71%:%
%:%134=72%:%
%:%135=73%:%
%:%136=74%:%
%:%137=75%:%
%:%138=76%:%
%:%139=77%:%
%:%140=78%:%
%:%141=79%:%
%:%142=80%:%
%:%143=81%:%
%:%144=82%:%
%:%145=83%:%
%:%146=84%:%
%:%147=85%:%
%:%148=86%:%
%:%149=87%:%
%:%150=88%:%
%:%151=89%:%
%:%152=90%:%
%:%153=91%:%
%:%154=92%:%
%:%155=93%:%
%:%156=94%:%
%:%157=95%:%
%:%158=96%:%
%:%159=97%:%
%:%160=98%:%
%:%161=99%:%
%:%162=100%:%
%:%163=101%:%
%:%164=102%:%
%:%165=103%:%
%:%166=104%:%
%:%167=105%:%
%:%168=106%:%
%:%169=107%:%
%:%170=108%:%
%:%171=109%:%
%:%172=110%:%
%:%173=111%:%
%:%174=112%:%
%:%175=113%:%
%:%176=114%:%
%:%177=115%:%
%:%178=116%:%
%:%179=117%:%
%:%180=118%:%
%:%181=119%:%
%:%182=120%:%
%:%183=121%:%
%:%184=122%:%
%:%185=123%:%
%:%186=124%:%
%:%187=125%:%
%:%188=126%:%
%:%189=127%:%
%:%190=128%:%
%:%191=129%:%
%:%192=130%:%
%:%193=131%:%
%:%194=132%:%
%:%195=133%:%
%:%204=137%:%
%:%208=139%:%
%:%220=141%:%
%:%221=142%:%
%:%222=143%:%
%:%223=144%:%
%:%224=145%:%
%:%225=146%:%
%:%226=147%:%
%:%227=148%:%
%:%228=149%:%
%:%229=150%:%
%:%230=151%:%
%:%231=152%:%
%:%232=153%:%
%:%233=154%:%
%:%234=155%:%
%:%235=156%:%
%:%236=157%:%
%:%237=158%:%
%:%238=159%:%
%:%239=160%:%
%:%240=161%:%
%:%241=162%:%
%:%242=163%:%
%:%243=164%:%
%:%244=165%:%
%:%245=166%:%
%:%254=168%:%
%:%266=170%:%
%:%267=171%:%
%:%268=172%:%
%:%269=173%:%
%:%270=174%:%
%:%271=175%:%
%:%272=176%:%
%:%273=177%:%
%:%274=178%:%
%:%275=179%:%
%:%276=180%:%
%:%277=181%:%
%:%278=182%:%
%:%279=183%:%
%:%280=184%:%
%:%281=185%:%
%:%282=186%:%
%:%283=187%:%
%:%284=188%:%

>>>>

%
\begin{isabellebody}%
\setisabellecontext{paper}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Introduction%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
As artifical reasoners become increasingly powerful, computers become capable of
performing complex ethical reasoning. Moreover, as regulators and academics call for ``ethical AI,"
the field of machine ethics becomes increasingly popular. Much prior work on machine ethics implements
relatively simple philosophical theories, partially as an artifact of this work emerging out of
Computer Science or Mathematics departments \cite{mesurvey}. Such work rarely capitalizes on the 
centuries of philosophical debate discussing the ethical theories in question. In this thesis, I attempt
to automate Kantian ethics, a sophisticated ethical theory, while staying faithful to philosophical
literature on the subject.

Automating a sophisticated ethical theory is interesting for two reasons. First, the proliferation of 
artifically autonomous agents is creating and will continue to create a demand for 
automated ethics. These agents must be able to reason about complex ethical theories 
that withstand philosophical scrutiny. Second, just as automated mathematical reasoning is a tool
for mathematicians, automated ethical reasoning is a tool that philosophers can use 
when reasoning about ethics. I argue that computational ethics can serve as another tool in a 
philosopher's toolbox, like a thought experiment or counterexample.

Modelling ethics without sacrificing the intracacies of an ethical theory is a 
challenging computational and philosophical problem. Simple and intuitive computational approaches, 
such as encoding ethics as a constraint satisfaction problem, fail to capture
the complexity of most philosophically plausible systems. Constraint satisfaction systems often default to some version of utiliarianism, 
the principle of doing the most good for the most people. Alternatively, they model basic moral 
principles such as ``do not kill," without modelling the theory that these principles originated from.
Modelling a more complex ethical theory will not only enable smarter philosophical machines, it will
also empower philosophers to study more complex ethical issues with the computer's help. The entire
field of philosophy is devoted to developing and testing robust ethical theories. Plausible machine
ethics must draw on plausible moral philosophy. Despite the importance of formalizing complex moral theories,
it is not immediately clear how to formalize complex theories like virtue ethics.

Kantian ethics, often described as ``formal" due to its rigid, rule-based structure, has been often 
floated as an attractive theory to automate \cite{powers, BL, lin}. This project's objective is to 
automate Kantian ethics. Roughly, my approach is to represent Kantian
ethics as an axioms in Carmo and Jones' Dyadic Deontic Logic (DDL), a modal logic designed to reason
about concepts like obligation \cite{CJDDL}. Such a representation of Kantian ethics in a logic
is called a ``formalization." I then embed this DDL formalization in higher-order logic 
(HOL) and implement it in the Isabelle/HOL theorem prover  \cite{isabelle}. I can then use Isabelle 
to automatically prove or disprove theorems (such as, ``murder is wrong") in my custom formalization, generating results
derived from the categorical imperative. Essentially, the computer is performing ethical reasoning using
the framework of Kantian ethics.

First, I recreate prior work implementing Dyadic Deontic Logic in Isabelle/HOL. Next, I present two 
prior attempts to formalize Kantian ethics in modal logic. The first is a naive
interpretation of Kant's categorical imperative that collapses to the base logic itself. The second is 
Moshe Kroy's partial formalization of the categorical imperative. As part of this presentation, I also
create a testing framework that can be used to evaluate these formalizations. In effect, my tests formalize
expected properties of the categorical imperative (such as the fact that it prohibits murder) and test 
whether or not these properties hold in a given formalization. My testing framework offers a method for
evaluating different formalizations against moral intuitions and philosophical literature. Second, based on
lessons learned from these prior formalization attempts, I contribute my own custom formalization of
the categorical imperative. My testing framework demonstrates how my custom formalization improves upon
the prior work. Third, I apply my system to an ethical dilemma to demonstrate its potential use. I also
consider the philosophical implications and value of automated ethics, studying questions about the 
possibility and responsibility of automated ethics.

I contribute implementations of two different interpretations of the categorical imperative, 
examples of how each implementation can be used to model and solve ethical scenarios, and tests that
examine ethical and logical properties of the system, including logical consistency, consistency
of obligation, and possibility of permissibility. I also contribute a logical formalization of the categorical
imperative that improves on previous work, an implementation of this formalization, and evidence of 
its improvement. Lastly, I demonstrate how such a system could be applied. The implementations themselves are usable models 
of ethical principles and the tests represent the kind of philosophical work that formalized ethics 
can contribute.

The rest of this project is structured as follows. In Chapter 1, I introduce the logical, computational, and 
philosophical frameworks underlying my project. In Chapter 2, I implement and test prior formalizations of the categorical
imperative. In Chapter 3, I contribute my own custom formalization and test it. Finally, in Chapter 4, 
I apply my custom formalization to an example ethical dilemma.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
In the rest of this chapter, I present the background necessary to undersand my work. The goal 
of this project is to automate sophisticated ethical reasoning. This requires three
components. First, the choice of an ethical theory that is both intuitively attractive and 
lends itself to formalization. Second, the choice of formal logic to model the theory in. Third, the 
choice of automation engine to implement the formal model in. Section \ref{sec:kant} introduces Kantian
ethics, Section \ref{sec:DDL} explains Carmo and Jones's Dyadic Deontic Logic \cite{CJDDL} as a 
base logic, and Section \ref{sec:isabelle} presents the Isabelle/HOL implementation of the logic.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Kantian Ethics \label{sec:kant}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Kantian ethics is an attractive choice of ethical theory to automate. Kant's writings inspired much of Western analytic philosophy. 
In addition to the rich neo-Kantian program, almost all major philosophical traditions after Kant have
engaged with his work. Much of Western libertarian political thought is inspired by Kant's deontology,
and his ethics have bled into household ethical thought. Deontology is seen as one of the three major 
schools of Western analytic ethics.

@@

For more of fine-grained properties of DDL see \cite{CJDDL} or this project's source code. DDL is a heavy logic and contains modal operators 
that aren't necessary for my analysis. While this expressivity is powerful, it may also cause performance
impacts. DDL has a large set of axioms involving quantification over complex higher-order
logical expressions. Proofs involving these axioms will be computationally expensive.  Benzmueller 
and Parent warned me that this may become a problem if Isabelle's automated proof tools begin to time out.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/Desktop/cs91r/paper/paper.thy%:%
%:%24=6%:%
%:%36=8%:%
%:%37=9%:%
%:%38=10%:%
%:%39=11%:%
%:%40=12%:%
%:%41=13%:%
%:%42=14%:%
%:%43=15%:%
%:%44=16%:%
%:%45=17%:%
%:%46=18%:%
%:%47=19%:%
%:%48=20%:%
%:%49=21%:%
%:%50=22%:%
%:%51=23%:%
%:%52=24%:%
%:%53=25%:%
%:%54=26%:%
%:%55=27%:%
%:%56=28%:%
%:%57=29%:%
%:%58=30%:%
%:%59=31%:%
%:%60=32%:%
%:%61=33%:%
%:%62=34%:%
%:%63=35%:%
%:%64=36%:%
%:%65=37%:%
%:%66=38%:%
%:%67=39%:%
%:%68=40%:%
%:%69=41%:%
%:%70=42%:%
%:%71=43%:%
%:%72=44%:%
%:%73=45%:%
%:%74=46%:%
%:%75=47%:%
%:%76=48%:%
%:%77=49%:%
%:%78=50%:%
%:%79=51%:%
%:%80=52%:%
%:%81=53%:%
%:%82=54%:%
%:%83=55%:%
%:%84=56%:%
%:%85=57%:%
%:%86=58%:%
%:%87=59%:%
%:%88=60%:%
%:%89=61%:%
%:%90=62%:%
%:%91=63%:%
%:%92=64%:%
%:%93=65%:%
%:%94=66%:%
%:%95=67%:%
%:%96=68%:%
%:%97=69%:%
%:%98=70%:%
%:%99=71%:%
%:%100=72%:%
%:%101=73%:%
%:%102=74%:%
%:%106=77%:%
%:%107=78%:%
%:%108=79%:%
%:%109=80%:%
%:%110=81%:%
%:%111=82%:%
%:%112=83%:%
%:%121=85%:%
%:%133=87%:%
%:%134=88%:%
%:%135=89%:%
%:%136=90%:%
%:%137=91%:%
%:%138=92%:%
%:%139=93%:%
%:%140=94%:%
%:%141=95%:%
%:%142=96%:%
%:%143=97%:%
%:%144=98%:%
%:%145=99%:%
%:%146=100%:%
%:%147=101%:%
%:%148=102%:%
%:%149=103%:%
%:%150=104%:%
%:%151=105%:%
%:%152=106%:%
%:%153=107%:%
%:%154=108%:%
%:%155=109%:%
%:%156=110%:%
%:%157=111%:%
%:%158=112%:%
%:%159=113%:%
%:%160=114%:%
%:%161=115%:%
%:%162=116%:%
%:%163=117%:%
%:%164=118%:%
%:%165=119%:%
%:%166=120%:%
%:%167=121%:%
%:%168=122%:%
%:%169=123%:%
%:%170=124%:%
%:%171=125%:%
%:%172=126%:%
%:%173=127%:%
%:%174=128%:%
%:%175=129%:%
%:%176=130%:%
%:%177=131%:%
%:%178=132%:%
%:%179=133%:%
%:%180=134%:%
%:%181=135%:%
%:%182=136%:%
%:%183=137%:%
%:%184=138%:%
%:%185=139%:%
%:%186=140%:%
%:%187=141%:%
%:%188=142%:%
%:%189=143%:%
%:%190=144%:%
%:%191=145%:%
%:%192=146%:%
%:%193=147%:%
%:%194=148%:%
%:%195=149%:%
%:%196=150%:%
%:%197=151%:%
%:%198=152%:%
%:%199=153%:%
%:%200=154%:%
%:%201=155%:%
%:%202=156%:%
%:%203=157%:%
%:%212=161%:%
%:%216=163%:%
%:%228=165%:%
%:%229=166%:%
%:%230=167%:%
%:%231=168%:%
%:%232=169%:%
%:%233=170%:%
%:%234=171%:%
%:%235=172%:%
%:%236=173%:%
%:%237=174%:%
%:%238=175%:%
%:%239=176%:%
%:%240=177%:%
%:%241=178%:%
%:%242=179%:%
%:%243=180%:%
%:%244=181%:%
%:%245=182%:%
%:%246=183%:%
%:%247=184%:%
%:%248=185%:%
%:%249=186%:%
%:%250=187%:%
%:%251=188%:%
%:%252=189%:%
%:%253=190%:%
%:%262=192%:%
%:%274=194%:%
%:%275=195%:%
%:%276=196%:%
%:%277=197%:%
%:%278=198%:%
%:%279=199%:%
%:%280=200%:%
%:%281=201%:%
%:%282=202%:%
%:%283=203%:%
%:%284=204%:%
%:%285=205%:%
%:%286=206%:%
%:%287=207%:%
%:%288=208%:%
%:%289=209%:%
%:%290=210%:%
%:%291=211%:%
%:%292=212%:%

### 2021-11-07 paper/output/document/upshot.tex
idk man

It is likely that, regardless of the strengths of the human-in-the-loop model, fully automated AI 
agents will exist. Even if developing this kind of AI is irresponsible,
such developments are likely and will require ethics engines, or risk no consideration of ethics at all. Even if 
fully automated AI is scary, such AI with automated ethics is better than such AI without. 
In such a world, the input parser in my ethics engine would have to be automated. This would require 
that the parser translate the AI agent's internal representation to the appropriate logical representation.
The input parser would need enough common sense reasoning to determine what circumstances are morally 
relevant to a maxim. This is a question that, like all of ethics, philosophers debate robustly
\footnote{Powers \cite{powers} identifies this as a challenge for automating Kantian ethics and briefly sketches 
solutions from O'Neill \cite{constofreason}, Silber \cite{silber}, and Rawls \cite{rawlsconstructivism}. }. 
It is likely that, just as different implementations of automated ethics choose 
a particular ethical theory and implement it, different implementations of such an input parser would 
need to adopt different interpretations of commonsense reasoning and morally relevant circumstances.
Automating this level of commonsense reasoning would represent signifiant technical progress in 
computational ethics.

@@

@@

Another, related benefit of computational ethics is that it makes formal ethics far less tedious. 
Certain subfields, such as philosophy of language, see such benefit in precision that they already 
use symbolic logic to represent philosophical concepts, just as mathematicians use symbolic logic to 
represent mathematical concepts. Some of this work requres tedious pencil and paper proofs to prove 
theorems, even when many of these theorems may not generate relevant philosophical insights. Ineractive 
theorem provers make formal logic more accessible. Isabelle can complete a proof, starting from first principles, 
in a matter of seconds that would take a logician pages to complete. Similarly, Nitpick can generate 
examples or counterexamples to a proposition in a brute force manner. The computer can generate hypotheses
that philosophers can then think about, just as it did for me to prompt the insight above.

@@

%:%467=383%:%
%:%479=385%:%
%:%480=386%:%
%:%481=387%:%
%:%482=388%:%
%:%483=389%:%
%:%484=390%:%
%:%485=391%:%

@@

%:%533=433%:%
%:%545=436%:%
%:%546=437%:%
%:%547=438%:%
%:%548=439%:%
%:%549=440%:%
%:%550=441%:%
%:%551=442%:%

@@

>>>>

It is likely that, regardless of the strengths of the human-in-the-loop model, fully automated AI 
agents will exist. Even if developing this kind of AI is irresponsible,
such developments are likely and will require ethics engines, or risk no consideration of ethics at all. Even if 
fully automated AI is scary, such AI with automated ethics is better than such AI without. 
In such a world, the input parser in my ethics engine would have to be automated. This would require 
that the parser translate the AI agent's internal representation to the appropriate logical representation.
The input parser would need enough common sense reasoning to determine what circumstances are morally 
relevant to a maxim. This is a question that, like all of ethics, philosophers debate robustly
\footnote{\citet{powers} identifies this as a challenge for automating Kantian ethics and briefly sketches 
solutions from \citet{constofreason}, \citet{silber}, and \citet{rawlsconstructivism}. }. 
It is likely that, just as different implementations of automated ethics choose 
a particular ethical theory and implement it, different implementations of such an input parser would 
need to adopt different interpretations of commonsense reasoning and morally relevant circumstances.
Automating this level of commonsense reasoning would represent signifiant technical progress in 
computational ethics.

@@

Below, I document the process used to arrive at the logical insight that the FUL is inconsistent if it
holds for maxims in which $(c \longrightarrow a) \wedge (c \longrightarrow g)$. For those uninterested
in the details of this exploration, it suffices to understand that I used Isabelle to show that if the 
FUL holds for badly formed maxims, then it is inconsistent. After I realized that my formalization was 
inconsistent, I made many failed attempts to diagnose and fix the problem before reaizing that the problem 
lay in badly formed maxims.

@@

Another, related benefit of computational ethics is that it makes formal ethics far less tedious. 
Certain subfields, such as philosophy of language, see such benefit in precision that they already 
use symbolic logic to represent philosophical concepts, just as mathematicians use symbolic logic to 
represent mathematical concepts. Some of this work requres tedious pencil and paper proofs to prove 
theorems, even when many of these theorems may not generate relevant philosophical insights. Interactive 
theorem provers make formal logic more accessible. Isabelle can complete a proof, starting from first principles, 
in a matter of seconds that would take a logician pages to complete. Similarly, Nitpick can generate 
examples or counterexamples to a proposition in a brute force manner. The computer can generate hypotheses
that philosophers can then think about, just as it did for me to prompt the insight above.

@@

%:%459=380%:%
%:%460=381%:%
%:%461=382%:%
%:%462=383%:%
%:%463=384%:%
%:%464=385%:%
%:%465=386%:%
%:%474=390%:%

@@

%:%525=431%:%
%:%526=432%:%
%:%527=433%:%
%:%528=434%:%
%:%529=435%:%
%:%530=436%:%
%:%531=437%:%
%:%540=440%:%

@@

%:%566=457%:%
%:%567=458%:%
%:%568=459%:%
%:%569=460%:%
%:%570=461%:%
%:%571=462%:%
%:%572=463%:%
